---
title: "Afternoon Session"
format: html
editor: visual
---

## Afternoon Session

```{r}
library(tidyverse)
library(palmerpenguins)
```

## Date-Time Info

ymd: if date is year-month-day

mdy: if date is month-day-year

dmy: if date is day-month-year

```{r}
lubridate::mdy("03-15-1998")

my_date_1 <- lubridate::mdy("11/7/2003")
class(my_date_1)

my_date_2 <- lubridate::ymd("19731025")

my_date_3 <- lubridate::dmy("08-Jun-1978")
class(my_date_3)

```

```{r}
time_1 <- lubridate::ymd_hm("2022-08-12 11:18")

```

Parse June 2nd 1982 at 4pm at UTC as a date in R

```{r}
time_2 <- lubridate::ymd_hm("1982-06-02 16:00", tz = "UTC")
```

Convert time zones with **with_tz()**

```{r}
lubridate::with_tz(time_2, "America/Los_Angeles")
```

```{r}
test_time <- ymd_hm("2022-12-10 4:32")
with_tz(test_time, "Europe/Monaco")

now<- with_tz(Sys.time(),"America/Los_Angeles")
now
```

Access different parts of day-time

```{r}
lubridate::day(now)
lubridate::week(now)
lubridate::hour(now)
lubridate::minute(now)


```

Data Frame:

```{r}
urchin_counts <- data.frame(
  date = c("Oct 12,2004", "Oct 18, 2007", "Feb 12, 2003", "Feb 14, 2008"),
  species = c("purple", "red", "red", "red"),
  size_mm = c(19,24,18,11)
  
)
# class(urchin_counts$date). returns character


```

```{r}
urchin_mean_by_month <- urchin_counts %>% 
  mutate(date = lubridate::mdy(date)) %>% 
  mutate(month = lubridate::month(date)) %>% 
  group_by(month) %>% 
  summarize(urchin_monthly_means = mean(size_mm))
```

## String with Stringr

Use str_detect to detect a string pattern

```{r}
my_str <- "The most common string to wirte is hello world"
str_detect(my_str,pattern = "hello world")
```

```{r}
my_str_2 <- c("burrito", "fist taco", "Taco salad", "super-burrito")
str_detect(my_str_2, "taco")
```

```{r}
starwars %>% 
  filter(str_detect(name, pattern = "Skywalker"))

starwars %>% 
  mutate(name_new = str_replace(name, pattern = "Sky" ,replacement = "Fire"))
```

str_trim to get rid of leading and trailing white space

srt_squish to get rid of interior white space

```{r}
my_string_3 <- "     Hello. My  name    is Teddy!   "
str_trim(my_string_3)
str_squish(my_string_3)
```

str_to_lower() makes the while string lowercase

str_to_upper() make the whole string CAPS

str_to_title() makes the first letter of each word Capitalized

str_to_sentence() makes the string return in sentence case

```{r}
starwars %>% 
  mutate(species_lower = str_to_lower(species))
```
